Is high social class always beneficial for survival? A study of northern Sweden 1801--2013 \footnote{Short title, "Is high social class beneficial for survival?"} \footnote{Funded by the Swedish Research Council under grants 349-2008-6592 and 421-2012-892.}

### 2018-10-09 popstud_open.Rmd
Small changes

The data for the present study come from two large population databases at the Demographic Data Base (DDB),Umeå University (http://www.cedar.umu.se), which provide us with micro-data for the Skellefteå and Umeå regions in northern Sweden (Demografiska databasen, CEDAR, Umeå universitet (2015). U15014. http://dx.doi.org/10.17197/U15014). The period 1801--1950 is covered by the database Poplink [@wee16]. Poplink is based on linked parish records, allowing us to reconstruct life biographies on people as long as they remained in the region. The records are linked within, but not
between, the regions.

@@

Figure \@ref(fig:plottaalla) shows the distribution
(per cent) of exposure times according to class for women and men (see also Tables \@ref(tab:womenexp) and \@ref(tab:menexp), respectively, in the Appendix). The farmer category was dominant until the middle of the 20th centuries but have become marginal in the latest periods, particularly among women. The category of workers is fairly stable over time. The highest class has increased substantially, albeit from a very low level
in the 19th century. The increase in academics in connection to the
establishment of Umeå University explains much of this increase in later time. Furthermore,
we observe a rather high percentage of missing social class in the early 19th century.
For men the proportion is somewhat higher in the age group 65 and above, those who were retired,
but still low enough to be used in our analyses (results not shown). If we disregard the very first period,
however, there were very few with occupation missing until 1950.

>>>>

The data for the present study come from two large population databases at the Demographic Data Base (DDB),Umeå University (http://www.cedar.umu.se), which provide us with micro-data for the Skellefteå and Umeå regions in northern Sweden [@ddb15]. The period 1801--1950 is covered by the database Poplink [@wee16]. Poplink is based on linked parish records, allowing us to reconstruct life biographies on people as long as they remained in the region. The records are linked within, but not
between, the regions.

@@

Figure \@ref(fig:plottaalla) shows the distribution
(per cent) of exposure times according to class for women and men (see also Tables \@ref(tab:womenexp) and \@ref(tab:menexp), respectively, as well as Table \@ref(tab:womendeaths), and \@ref(tab:mendeaths) for number of deaths during the different periods in the Appendix). The farmer category was dominant until the middle of the 20th centuries but have become marginal in the latest periods, particularly among women. The category of workers is fairly stable over time. The highest class has increased substantially, albeit from a very low level
in the 19th century. The increase in academics in connection to the
establishment of Umeå University explains much of this increase in later time. Furthermore,
we observe a rather high percentage of missing social class in the early 19th century.
For men the proportion is somewhat higher in the age group 65 and above, those who were retired,
but still low enough to be used in our analyses (results not shown). If we disregard the very first period,
however, there were very few with occupation missing until 1950.
